---
title: "rmarkdown_day2modeling"
author: "Lea Briard"
date: "May 16, 2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, results = T, warning = F)
```


```{r}
## packages ##
library(tidyverse)
library(magrittr)
library(GGally)
library(ggplot2)
library(plotly)


#task 1 using 
setwd("C:/Users/workshop/Desktop")

#load RData version of the file
load("C:/Users/workshop/Desktop/ld_pop_prism.RData")

# load csv version of the dataset
datacsv<-read_csv("C:/Users/workshop/Desktop/ld_pop_prism.csv")
```

#task 2: 
use ggpairs to obtain a 4 by 4 summary plot of precipitation, average temperature, 
pop size, number of Lyme disease cases
```{r}
colnames(fulldata)
ggpairs(fulldata, columns = c("prcp", "avtemp", "size", "cases") )
```

#task 3
create 2 new columns for log10(size) and log10(cases+1)

```{r}
fulldata%<>% 
  mutate(log10_size = log10(size)) %>%
  mutate(log10_cases = log10(cases+1))

ggpairs(fulldata, columns = c("prcp", "avtemp", "log10_size", "log10_cases"))
## why do we add 1 to the number of cases? because log(0) is undefined (not a real number) 
```


# linear model #
# task 4
```{r}
set.seed(222); fulldata_100<- fulldata %>% sample_n(100) #1st call
plot_prcp_temp <- fulldata_100 %>%
  ggplot(aes(x=prcp, y=avtemp))+geom_point()
plot_prcp_temp + geom_smooth(aes(x=prcp, y=avtemp),method="lm")
```

#Task 6: 
```{r}
#Create a linear model (lm) object with a call like myModel <- lm(y ~ x, data = myData) for the
#subsetted data, where y=avtemp and x=prcp. In addition, view the summary with a call along the lines of
#summary(myModel)

model_prcptemp<-lm(prcp~avtemp, fulldata_100)
summary(model_prcptemp)
summary(model_prcptemp)$coefficients[2,1]
summary(model_prcptemp)$coefficients[2,4]
```


# Task 7: 

What is the slope of the line you plotted in Task 5, and is the slope significantly different from 0 (p<0.05)?
```{r}
#summary(model_prcptemp)$coefficients[2,1]
#[1]29.68683
#summary(model_prcptemp)$coefficients[2,4]
#[1] 3.190341e-06
```

#modelr package#
#task 8: 
Write a single line of code to generate a ggplot of total population size by year.

```{r}
fulldata %>% group_by(year) %>%
  summarise(popsize_per_year=sum(size))%>%
  ggplot(aes(x=year, y=popsize_per_year))+geom_point()
``` 

# Grouped data frames versus nested data frames
#Task 9: 
Create a data frame called "by_state" from the main data frame, that groups by state, and inspect it.
```{r}
by_state<-fulldata%>% group_by (state)
by_state
##Task 10: Next, update this new data frame so that it is nested (simply pass it to nest). Again, inspect the
##data frame by typing its name in the console so see how things changed.
by_state %<>% nest
by_state

by_state$data[[10]]
```

#task 12 :
Write a function that takes a data frame as its argument and returns a linear model object that predicts size by year.

```{r}
linearmodel <- function(df){
  lm(size~year,data=df)
}
```
# introduction to the functions of purrr
```{r}
models <- purrr::map(by_state$data, linearmodel)
```
Function conflicts: why the purrr::?
to avoid conflicts and make sure that the function we use comes from the appropriate package purrr apply the function linearmodel on each element of by_state


#Task 13: 
Add a column to the by_state dataframe, where each row (state) has its own model object.

```{r}
by_state%<>%
  mutate(model= purrr::map(data, linearmodel))

library(modelr)
by_state%<>%
  mutate(resids= purrr::map2(data, model, add_residuals))
```
#Task 14: 
Run these commands and inspect "resids". What is the structure of "resids"?

```{r}
str(by_state$resids)
## resids is nested in by_state (same level as data and model)
```


#Task 15: 
Write a function that accepts an object of the type in the resids list, and returns a sum of the
absolute values, i.e. ignoring sign: abs(3)+abs(-2)=5. Use the function to add a column called totalResid
to by_state that provides the total size of residuals summed over counties and years.

```{r}
#x is the dataframe by_state
sum_resids <- function (x) {
  sum(abs(x$resid))
 }

by_state%<>%
  mutate(totalResid = purrr::map (resids, sum_resids))
```

#Task 16: 
Write a function that accepts a linear model and returns the slope (model M has slope M$coefficients[2]) and then use this function to create a new column called slope in the by_state dataframe, that is the slope for each state.

```{r}
# x is by_state$model
slope <- function (x){
  sl<-x$coefficients[2]
  return(sl)
}

by_state%<>%
  mutate(slope = purrr::map (model, slope))


slopes <- unnest(by_state, slope)
totalResids <- unnest(by_state, totalResid)
```

#Task 17: 
Plot the growth rate (slope value) for all states.

```{r}
plot_growth_bystate<- slopes %>%
  ggplot(aes(x=state, y=slope))+geom_point()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
plot_growth_bystate
```

#Task 18: 
Plot the total resisduals for all states.
```{r}
plot_growth_bystate<- totalResids %>%
  ggplot(aes(x=state, y=totalResid))+geom_point()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
plot_growth_bystate
```


#Task 19: 
Repeat Tasks 9 and 10 using a different data frame name, by_state2.

```{r}
#Grouped data frames versus nested data frames
#Task 9: Create a data frame called "by_state" from the main data frame, that groups by state, #and inspect it.

by_state2<-fulldata%>% group_by (state)
by_state2

##Task 10: Next, update this new data frame so that it is nested (simply pass it to nest). Again, inspect the
##data frame by typing its name in the console so see how things changed.
by_state2 %<>% nest
by_state2

by_state2$data[[10]]
```

#Task 20: 
Write a function that accepts an element of the by_state2$data list-column and returns the spearman correlation coefficient between Lyme disease cases and precipitation
```{r}
#df is by_state2
Corr<- function (df) {
  suppressWarnings(cor.test(df$cases, df$prcp, method="spearman")$estimate)
}

by_state2%<>%
  mutate(corr_coef = purrr:: map (data, Corr))
spcor<- unnest (by_state2, corr_coef)
spcor%<>% arrange(desc(corr_coef))
spcor$state <- factor(spcor$state, levels=unique(spcor$state))
myplot<-ggplot(spcor,aes(state,corr_coef))+geom_point()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))



```

